### 2014-11-05 contract-language.tex
edits

The syntax of our core contract language is shown in Fig.
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed effect variables. We use a
special effect variable ($\cureff$) to denote the effect of \emph{current
operation} - the operation for which a contract is being written. The type of
an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
the effect. We admit disjuntion in types to let an effect variable range over
multiple operation names. The contract $\small \forall (a : \tau_1 \vee
\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
\vee \oper{a}{\tau_2}) \Rightarrow \psi$.

@@

\name contracts are constraints over axiomatic definitions of program
executions.Fig.~\ref{sem:contracts} summarizes artifacts relevant to define an
axiomatic execution. We formalize an axiomatic execution as a tuple $\Exec$,
where $\EffSoup$, called the \emph{effect soup}, is the set of all effects
generated during the program execution, and $\visZ,\soZ,\sameobjZ \subseteq
\EffSoup \times \EffSoup$ are \emph{visibility}, \emph{session order}, and
\emph{same object} relations, respectively, witnessed over generated effects at
run-time.

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{c}{
  {op} \in \mathtt{Operation} \qquad
  \tau \in \mathtt{Consistency\; Class}
}\\
\cv(\tau) & \in & \mathtt{Store\; Contract\; of \; \tau} & \coloneqq & \scc,
  \ccc, \ecc\\
{\sigma} & \in & \mathtt{Session} & \coloneqq & \cdot \ALT (op,\tau); \sigma \\
\Sigma 	& \in & \mathtt{Session\;Soup}   & \coloneqq &
  \langle s,{\sigma} \rangle \pll \Sigma \ALT \emptyset \\
				&			&	\mathtt{Config}		  & \coloneqq & \E;\Sigma \\
\end{array}
\end{smathpar}

We model the system as a tuple $\E;\Sigma$, where the axiomatic execution $\E$
captures the data store's current state, and session soup $\Sigma$ is the
set of concurrent client sessions interacting with the store. A session is
modeled simply as a sequence of replicated data type operations
($\mathit{op}$), tagged with the consistency class ($\tau$) of their contracts
(as determined by the contract classification scheme). We assume a rewrite
relation of form:

\vspace{-1.7em}
\begin{smathpar}
  \auxred{} {\E;\langle (op,\tau);\sigma \rangle \pll \Sigma} {\eff}
    {\E';\langle \sigma \rangle \pll \Sigma'}
\end{smathpar}

\noindent on the system state. The relation captures the progress of
the execution (from $\E$ to $\E'$)  due to the successful completion
of a client operation ($\mathit{op}$) from one of the sessions in $\Sigma$,
generating a new effect ($\eff$). If the resultant execution ($\E'$)
satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
\models \cv(\tau)$), then we say that the store has \emph{enforced} the
contract $\cv(\tau)$ in the execution $\E'$.

With help of the axiomatic model, we now state the soundness of
contract enforcement as follows:

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
and let $\tau$ denote the consistency class of $\cv$ as determined by
the contract classification scheme. For all well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
 {\E';\langle \sigma \rangle \pll \Sigma}$,
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
\end{theorem}

\noindent The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
store, extended with the classification scheme shown in
Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
proof of the theorem is given in a tech report and is provided as part of the
supplementary material~\cite{techrep}\footnote{The supplementary material and
tech report also defines an operational semantics that concretizes the rewrite
relation ($\xrightarrow{}$), along with the proof that it correctly enforces
all \name contracts.}.

>>>>

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed effect variables. We use a
special effect variable ($\cureff$) to denote the effect of \emph{current
operation} - the operation for which a contract is being written. The type of
an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
the effect. We admit disjuntion in types to let an effect variable range over
multiple operation names. The contract $\small \forall (a : \tau_1 \vee
\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
\vee \oper{a}{\tau_2}) \Rightarrow \psi$.

@@

\name contracts are constraints over axiomatic definitions of program
executions. Figure~\ref{sem:contracts} summarizes artifacts relevant to define
an axiomatic execution. We formalize an axiomatic execution as a tuple $\Exec$,
where $\EffSoup$, called the \emph{effect soup}, is the set of all effects
generated during the program execution, and $\visZ,\soZ,\sameobjZ \subseteq
\EffSoup \times \EffSoup$ are \emph{visibility}, \emph{session order}, and
\emph{same object} relations, respectively, witnessed over generated effects at
run-time.

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

We model the system as a tuple $\E,\Sigma$, where the axiomatic execution $\E$
captures the data store's current state, and session soup $\Sigma$ is the set
of concurrent client sessions interacting with the store. A session $\sigma$ is
a sequence of pairs composed of replicated data type operations $\mathit{op}$,
tagged with the consistency class $\tau$ of their contracts (as determined by
the contract classification scheme). We assume a rewrite relation of form:

\vspace{-1.7em}
\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}

\noindent on the system state. The relation captures the progress of the
execution (from $\E$ to $\E'$)  due to the successful completion of a client
operation $\mathit{op}$ from one of the sessions in $\Sigma$, generating a new
effect $\eff$. If the resultant execution $\E'$ satisfies the store contract
$\cv_\tau$ (i.e., $\E \models \cv_\tau$), then we say that the store has
\emph{enforced} the contract $\cv_\tau$ in the execution $\E'$. With help of
the axiomatic model, we now state the soundness of contract enforcement as
follows:

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation
$\mathit{op}$, and let $\tau$ denote the consistency class of $\cv$ as
determined by the contract classification scheme. For all well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff} {\E', \sigma
\pll \Sigma}$, if $\E' \models \cv_\tau[\eff/\cureff]$, then $\E' \models
\cv[\eff/\cureff]$
\end{theorem}

\noindent The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
store, extended with the classification scheme shown in
Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
proof of the theorem is given in the supplementary material\footnote{The
supplementary material also defines an operational semantics that concretizes
the rewrite relation ($\xrightarrow{}$), along with the proof that it correctly
enforces all \name contracts.}.

### 2014-11-05 core-lang.tex
Edits
edits

Fig.~\ref{sem:oper} summarizes the artifacts needed. We formalize an
abstract model of a replicated data store as a tuple $\E;\Sigma$, where $\E$
denotes the execution state, and $\Sigma$, the session soup, is the set of
concurrent client sessions interacting with the store. The execution state
($\E$) of a data store is modeled as a tuple $\Exec$ where $\EffSoup$,
called the \emph{effect soup}, is a set of effects generated during the
execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
visibility and session order relations, respectively, witnessed over
generated effects. A session is modeled simply as a sequence of replicated
data type operations ($op$), tagged with the consistency class ($\tau$) of
their contracts (as determined by the contract classification scheme). Our
contract language and classification scheme is independent of store
semantics. Accordingly, we avoid making any specific assumptions about the
semantics of the store. However, we do assume the existence of a reduction
relation of the form:

>>>>

Figure~\ref{sem:oper} summarizes the artifacts needed. We formalize an abstract
model of a replicated data store as a tuple $\E;\Sigma$, where $\E$ denotes the
execution state, and $\Sigma$, the session soup, is the set of concurrent
client sessions interacting with the store. The execution state ($\E$) of a
data store is modeled as a tuple $\Exec$ where $\EffSoup$, called the
\emph{effect soup}, is a set of effects generated during the execution, and
$\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are visibility and session
order relations, respectively, witnessed over generated effects. A session is
modeled simply as a sequence of replicated data type operations ($op$), tagged
with the consistency class ($\tau$) of their contracts (as determined by the
contract classification scheme). Our contract language and classification
scheme is independent of store semantics. Accordingly, we avoid making any
specific assumptions about the semantics of the store. However, we do assume
the existence of a reduction relation of the form:

### 2014-11-05 motivation.tex
edits

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
exposes the contract construction language as a Haskell library}) that
axiomatically specify the set of allowed executions involving this operation.
In the case of the bank account, any execution that does not exhibit the
anomalies described in the previous section is a \emph{well-formed} execution
on the bank account object. By specifying the set of legal executions for each
data type in terms of a trace of operation invocations on that type, \name\
ensures that all executions over that type are well-formed.

>>>>

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} (Figure~\ref{fig:contract-lang}\footnote{\name
exposes the contract construction language as a Haskell library}) that
axiomatically specify the set of allowed executions involving this operation.
In the case of the bank account, any execution that does not exhibit the
anomalies described in the previous section is a \emph{well-formed} execution
on the bank account object. By specifying the set of legal executions for each
data type in terms of a trace of operation invocations on that type, \name\
ensures that all executions over that type are well-formed.
